The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-07-17 WavesData-2019/LatexNotes/notes.tex
NM

% Basic physical parameters and scales
\newcommand{\depth}{h}
\newcommand{\dup}{\depth_{-}}
\newcommand{\ddn}{\depth_{+}}
\newcommand{\freqp}{f_p}
\newcommand{\lam}{\lambda}
\newcommand{\lamup}{\lam_{-}}
\newcommand{\lamfac}{N}
\newcommand{\amp}{a}
%\newcommand{\etastd}{\eta_{\text std}}
% Statistical quantities
\newcommand{\skw}{\text{skew}}
\newcommand{\var}{\text{var}}
% Dimensionless parameters
\newcommand{\epsup}{\eps_0}
\newcommand{\delup}{\delta_0}
\newcommand{\drat}{D}
\newcommand{\dratdn}{\drat_0}
\newcommand{\ampp}{\mathcal{A}}
% Hamiltonian and Gibbs stuff
\newcommand{\Ham}{H}
\newcommand{\Hup}{\Ham^{-}}
\newcommand{\Hdn}{\Ham^{+}}
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\Gup}{\Gibbs^{-}}
\newcommand{\Gdn}{\Gibbs^{+}}
\newcommand{\meanup}[1]{\mean{#1}_{-}}
\newcommand{\meandn}[1]{\mean{#1}_{+}}

@@

We introduce the following characteristic scales
\begin{align}
\label{ampdefn}
&\amp = \mean{\eta^2}^{1/2} 
&&\mbox{\em characteristic wave amplitude} \\
\label{lamdefn}
&\lam = \sqrt{g \depth} / \freqp
&&\mbox{\em characteristic wave length}
\end{align}
where $\mean{\cdot}$ indicates the mean of a quantity. Note that the characteristic wavelength $\lam = \lam(x)$ takes different values upstream and downstream of the ADC. We remark that experimental measurements indicate that the characteristic amplitude $\amp = \mean{\eta^2}^{1/2}$ is nearly the same on both values of the ADC. Hence, we will not distinguish between upstream and downstream values of $\amp$.
We introduce the following dimensionless variables
\begin{align}
&u = \eta/\amp
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = \frac{2 \pi \xi}{\lamfac \lam} = \frac{2 \pi (x-ct)}{\lamfac \lam}
&&\mbox{\em dimensionless position in moving frame} \\
& \tilde{t} = t \freqp / \lamfac
&&\mbox{\em dimensionless time}
\end{align}
We will consider the physical domain $\xi \in [0, \lamfac \lam]$, which is an integer multiple of the characteristic wavelength where $\lamfac$ will be selected later (probably about 5). Above, we have chosen the scale to normalize position as $\lamfac \lam/(2\pi)$ so that the dimensionless domain can be taken (after a shift) to be $\tilde{x} \in [-\pi,\pi]$. Accordingly, we have also chosen the time scale as $\lamfac/\freqp$, so that a wave traveling at the local wavespeed $c = \lam \freqp = \sqrt{g \depth}$ crosses the physical domain $x \in [0, \lamfac \lam]$ in one unit of dimensionless time. Said differently, the dimensionless wave speed is $(2 \pi)^{-1}$ so that a wave crosses the domain $[-\pi,\pi]$ in unit time.

@@

Dropping the tildes, the dimensionless KdV equation takes the form
\begin{equation}
\label{dimlessKdV}
u_t + {3 \pi} \epsup \drat^{-1} \, u u_x + \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \drat \, u_{xxx} = 0
\qquad \text{for } x \in [-\pi,\pi]
\end{equation}
where dimensionless variables are as follows
\begin{align}
&\epsup = \amp/\dup \, , \\
&\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g} \, \\
&\drat = {\depth}/{\dup} \, .
\end{align}
Here, $\drat$ is the local dimensionless depth, which changes value crossing the ADC. In particular, since \eqref{dimlessKdV} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$. Then we have
\begin{equation}
\drat = 
\begin{cases}
1 		&\quad \mbox{for } t<0 \\
\dratdn 	&\quad \mbox{for } t>0
\end{cases}
\end{equation}
where $\dratdn = \ddn/\dup$.
See Table \ref{paramtable} for a summary of important parameters.

@@

>>>>

% Basic physical parameters and scales
\newcommand{\freqp}{f_p}
\newcommand{\amp}{a}
%\newcommand{\etastd}{\eta_{\text std}}
% Parameters that change crossing the ADC
\newcommand{\depth}{d}
\newcommand{\dup}{\depth_{-}}
\newcommand{\ddn}{\depth_{+}}
\newcommand{\dupdn}{\depth_{\pm}}
\newcommand{\lam}{\lambda}
\newcommand{\lamup}{\lam_{-}}
\newcommand{\lamdn}{\lam_{+}}
\newcommand{\lamupdn}{\lam_{\pm}}
\newcommand{\lamfac}{N}
\newcommand{\drat}{D}
\newcommand{\dratdn}{\drat_+}
\newcommand{\dratupdn}{\drat_{\pm}}
% Statistical quantities
\newcommand{\skw}{\text{skew}}
\newcommand{\var}{\text{var}}
\newcommand{\En}{\mathcal{E}}
% Dimensionless parameters
\newcommand{\epsup}{\eps_0}
\newcommand{\delup}{\delta_0}

\newcommand{\ampp}{\mathcal{A}}
% Hamiltonian and Gibbs stuff
\newcommand{\Ham}{H}
\newcommand{\Hup}{\Ham^{-}}
\newcommand{\Hdn}{\Ham^{+}}
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\Gup}{\Gibbs^{-}}
\newcommand{\Gdn}{\Gibbs^{+}}
\newcommand{\meanup}[1]{\mean{#1}_{-}}
\newcommand{\meandn}[1]{\mean{#1}_{+}}

@@

 \subsubsection{Most conventional nondimensionalization}
\label{nondim1}

We introduce the following characteristic scales
\begin{align}
\label{ampdefn}
&\amp = \mean{\eta^2}^{1/2} 
&&\mbox{\em characteristic wave amplitude} \\
\label{lamdefn}
&\lamupdn = \sqrt{g \dupdn} / \freqp
&&\mbox{\em characteristic wave length}
\end{align}
where $\mean{\cdot}$ indicates the mean of a quantity. Note that the characteristic wavelength $\lam = \lamupdn$ takes different values upstream and downstream of the ADC. We remark that experimental measurements indicate that the characteristic amplitude $\amp = \mean{\eta^2}^{1/2}$ is nearly the same on both values of the ADC. Hence, we will not distinguish between upstream and downstream values of $\amp$.
We introduce the following dimensionless variables
\begin{align}
&u = \eta/\amp
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = \frac{\pi \xi}{\lamfac \lam} = \frac{\pi (x-ct)}{\lamfac \lam}
&&\mbox{\em dimensionless position in moving frame} \\
& \tilde{t} = t \freqp / \lamfac
&&\mbox{\em dimensionless time}
\end{align}
We will consider the physical domain $\xi \in [-\lamfac \lam, \lamfac \lam]$, which is an integer multiple of the characteristic wavelength, where $\lamfac$ will be selected later (probably about 5). Above, we chosen to normalize position by the scale $L = \lamfac \lam/\pi$ so that the dimensionless domain is $\tilde{x} \in [-\pi,\pi]$. Accordingly, we have also chosen the time scale as $\lamfac/\freqp$, so that a wave traveling at the local wavespeed $c = \lam \freqp = \sqrt{g \depth}$ crosses the physical domain of length $2 \lamfac \lam$ in one unit of dimensionless time. Said differently, the dimensionless wave speed is $(2 \pi)^{-1}$ so that a wave crosses the domain $[-\pi,\pi]$ in unit time.

@@

Dropping the tildes, the dimensionless KdV equation takes the form
\begin{equation}
\label{dimlessKdV}
u_t + {3 \pi} \epsup \drat^{-1} \, u u_x + \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \drat \, u_{xxx} = 0
\qquad \text{for } x \in [-\pi,\pi]
\end{equation}
where dimensionless variables are as follows
\begin{align}
&\epsup = \amp/\dup \, , \\
&\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g} \, \\
&\drat = {\depth}/{\dup} \, .
\end{align}
Here, $\drat = \dratupdn$ is the local dimensionless depth, which changes value crossing the ADC. In particular, since \eqref{dimlessKdV} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$. Then we have
\begin{equation}
\drat = 
\begin{cases}
1 		&\quad \mbox{for } t<0 \\
\dratdn 	&\quad \mbox{for } t>0
\end{cases}
\end{equation}
where $\dratdn = \ddn/\dup$.
See Table \ref{paramtable} for a summary of important parameters.

@@

%----------------------------------------------------------------------------%
% Energy nondimensionalization
%----------------------------------------------------------------------------%
\subsubsection{Energy nondimensionalization}

Instead of scaling the surface displacement on the characteristic amplitude, a slightly different approach is to normalize the energy to unity. Consider waves in the domain $\xi \in [-\pi L, \pi L]$ where $L$ is a length-scale we define as
\begin{equation}
L = \lamfac \lam / \pi
\end{equation}
Consider the so-called energy of $\eta$
\begin{equation}
\En [\eta] = \frac{1}{2} \int_{-\pi L}^{\pi L} \eta^2 \dx = \pi L \mean{\eta^2} \equiv E
\end{equation}
We introduce the normalized variable
\begin{align}
&u = E^{-1/2}L^{1/2} \eta \\
&\tilde{x} = \xi/L
\end{align}
Then, in dimensionless variables, we have
\begin{equation}
\En [u] = \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = \pi \mean{u^2} = \frac{\pi L}{E} \mean{\eta^2} = 1
\end{equation}
That is, the energy is normalized to unity.

To compare to the previous formulation, we note that
\begin{align}
\amp \equiv \mean{\eta^2}^{1/2} = \sqrt{ \frac{E}{\pi L} }
\end{align}
So that the new normalization factor is
\begin{equation}
E^{1/2}L^{-1/2} = \sqrt{\pi} \amp
\end{equation}
In other words, the normalizations for surface displacement used here versus that in Section \ref{nondim}
 differ by a factor of $\sqrt{\pi}$. Fortunately, the normalization for position is identical in the two. We do have to remember that the normalization factor for position $L = \lamfac \lam / \pi$ takes two different values on either side of the ADC.

TO DO: How to scale time ....
